---
Title: "Classification"
Author:"Atmin Sheth, Andrew Sen "
output:
  html:
    df_print:paged
    pdf_document:default
editor: 
  chunck_output_type: inline

---

**Author:**
  Atmin Sheth
  Andrew Sen
  
**Date**
09-23-2022


**Data**

The note book is acquired on kragle  dataset called Airlines
showing the airline and their flights being delay(1) or not (0)

The predictor include:
*   id: int (unique number )
*   Airline: factor of char 
*   Flight: int (unique number)
*   AirpotFrom: char
*   AirportTo:  char
*   DayofWeek: int(constant)
*   Time: int
*   Delay: boolean Delay(1) not delay(0)
targeted predictor:
*   Airline
*   Flight
*   Delay

***overview of Logistic regression***
Logistic Regression allows to classify the data into n- parts  to determine the section the factors belong and predict upon that 
for this dataset is a classification of if the airlines will have delay or not , flight is to see a corelatrion wiht airline

**strengths of logistic **
The benefit of using logistic reggresuib model let's you separate the classes and see a distinct differences, it is easy computation 
where the result is in probability bases on the predictor and and relation of probability or mean to other predictors. The output
is easy to comprehend.

**weakness of logistic**
there is a chance of too close of probability so not able to classify due to being clustered.  Not able to sometimes 
make a model line. 

**Strength of naive bayes**
Naive bayes is easy to implement and interpret. It works well with small data. it also works well with high dimension

**weakness of naive bayes**
Naive bayes tend to outperform in larger set.there are many guesses that are not sometimes made in a train dataset.there is a limitation if the predictors are
not independent.

**the evaluation**
The benefit of using this was see the probability of a airline being delay or not.

##Reading csv file
want find the difference of delay and not delay
the airlines that are tend to be delay
```{r}
flights <- read.csv("Data/Airlines.csv", header = TRUE)
str(flights)
```

###data cleean
getting the data we want to work with that be airlines,flight and  delay


```{r}
flights <- flights[,c(2,3,9)]
flights$Airline <- as.numeric(factor(flights$Airline))
head(flights)
```

##Data Exploration
spliting the data to training and testinng in a ration of 80/20(.4) train/test


```{r}
set.seed(3)
i <- sample(1:nrow(flights), .4*nrow(flights),replace=FALSE)
train <- flights[-i,]
test <- flights[i,]
```

```{r}
summary(train)
```

seeing the head of the train set

```{r}
head(train)
```

the end of the train data set 
```{r}
tail(train)
```
Seeing how many airplines we are working with in training 


dimention
```{r}
dim(train)
```

```{r}
table(train$Airline)

```
Seeing the sum of delay flight 
how many of the files in the train dataset was delaye
```{r}
sum(train$Delay)
```
where are the flight most going to 
```{r}
dim(train )
```




bar graph to show the delay vs non delay in all of the airline in train data set
```{r}
count <- table(train$Delay)
bp<- barplot(count,main="Flights arival times", names.arg=c("Delay", "on Time"),cex.names=.8)
```

There are more delay in overall flights then reaching in time 




let's see a  airlines delay in a subset of about 100 in  train
```{r}
sub<-train[1:500,]
T<-table(sub$Airline,(sub$Delay==1))
plot(T, ylab="id delay", col="blue", main="Airlines Arival Delay or not")
```
 
in the subset it can see that there are airlines, w



##logistic regression predictor

```{r}
glm1 <- glm(Delay~. , family=binomial, data=train)
summary(glm1)
```
It took 4 itteration to get the results
There is a drop from the Null deviance to the Residual deviance
for all two predict or for airline and flight there is a good p value of less the 2e^-16 

##naive basis

```{r}
library(e1071)
nb1<- naiveBayes(Delay~.,data=train)
nb1
```
here we can see a mean of airlines and flights
the likelihood of airline and flights being delay 

#model testing 
```{r}
probs <- predict(glm1, newdata = test, type="response")
preds <- ifelse(probs==0,0,1)
acc1<- mean(preds==test$Delay)
print(paste("glm1 accuracy =",acc1))
table(preds,test$Delay)
print("prediction of naive bais")
predN <- predict(nb1,newdata = test, type="class")
table(predN,test$Delay)
print("mean of delays in the naive bais ")
mean(predN==test$Delay)

```

From this models of test data it can be seen that the rate of being late is low 
the mean shows the rate being .58 






